---
title: "3_transform_drinkers_data"
---

"This script transforms drinkers data to fit the normal distribution."

## Read in necessary packages, functions & data, & specify preferences

# Setup
```{r, warning = FALSE}

# Read in R packages
library(tidyverse)
library(readr)
library(tidyr)
library(dplyr)
library(labelled)
library(haven)
library(mice)
library(clipr)
library(forecast)
library(MASS)
library(ggplot2)

# Set default theme for plots
theme_set(theme_bw(base_size = 10))

# bias against scientific notiation
options(scipen=999)

# Read in drinkers data
drinkers <- readRDS("U:/SIMAH/SIMAH_workplace/nhis/intersectionality/cleaned_data/5_nhis_alc_clean_drinkers_only.RDS")
```

Review basic stats of estimated grams per day data for drinkers
```{r}
drinkers %>% summarise(min(alc_daily_g), max(alc_daily_g), mean(alc_daily_g))
drinkers %>% summarise(min(alc_daily_g_capped_200), max(alc_daily_g_capped_200), mean(alc_daily_g_capped_200))
```
# Transform consumption data for drinkers

Check recommended lambda with boxcox
```{r}
b <- MASS::boxcox(lm(drinkers$alc_daily_g_capped_200 ~ 1))
lambda <- b$x[which.max(b$y)] # 0.0606

lambda2 <- forecast::BoxCox.lambda(drinkers$alc_daily_g_capped_200)  # -0.054
```
As the suggested lambda are close to 0, log transformation is appropriate

Log transformation
```{r}
drinkers$capped_daily_grams_log <- log(drinkers$alc_daily_g_capped_200)

# Distribution plot 
ggplot(drinkers, aes(x=capped_daily_grams_log), y) + geom_histogram(bins=100) + 
ylim(0,20000)

# qq plot
qqnorm(drinkers$capped_daily_grams_log, pch = 1, frame = FALSE)
qqline(drinkers$capped_daily_grams_log, lwd = 2)
```
# Check transformation method by back-transforming and re-calculating the back transformed mean
```{r}
drinkers$test <- exp(drinkers$capped_daily_grams_log)
mean(drinkers$test)
```
Save log transformed data
```{r}
saveRDS(drinkers, "U:/SIMAH/SIMAH_workplace/nhis/intersectionality/cleaned_data/6_nhis_alc_clean_transformed_drinkers_log.RDS")
```
